---
title: "Analysis EMG"
author: '[Antonio Schettino](https://osf.io/zbv65/ "Antonio Schettino")'
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    highlight: tango
    theme: united
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

# ### install packages
# install.packages(c(
#   "knitr", "kableExtra",
#   "tidyverse",
#   "rcompanion", "ARTool", "ez"
#   "viridis", "cowplot"
# ),
# dependencies = TRUE
# )
# devtools::install_github("mikabr/ggpirate", dependencies = TRUE)

### load packages
library(knitr)
library(kableExtra)
library(tidyverse)
library(rcompanion)
library(ARTool)
library(ez)
library(viridis)
library(ggpirate)
library(cowplot)

# for RMarkdown
options(digits = 2) # number of decimal digits
opts_chunk$set(
  fig.dim = c(10, 10),
  warning = FALSE, # no package warnings
  message = FALSE # no package messages
)

# ggplot custom theme
theme_custom <- theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(
      hjust = .5,
      size = 20
    ),
    legend.title = element_blank(),
    plot.title = element_text(size = 26, hjust = .5),
    legend.box.background = element_rect(color = "transparent") # transparent outer line
  )

seed_smorfia <- 1 # Italia!
set.seed(seed_smorfia) # seed for random number generation

```

***
***

# Cue & Target

```{r all_data}

EMG <- read_csv("EMG.csv") %>%
  pivot_longer(
    Reappraisal_Unexpected:Appraisal_Expected,
    names_to = "condition",
    values_to = "amplitude"
  ) %>%
  separate(
    condition,
    c("ER_Strategy", "Expectation"),
    sep = "_",
    remove = TRUE
  ) %>%
  mutate(
    Participant = as_factor(Participant),
    Believed_Story = as_factor(Believed_Story),
    Stimulus = as_factor(Stimulus),
    ER_Strategy = as_factor(ER_Strategy),
    Expectation = as_factor(Expectation)
  )

# summary
EMG_summary <-
  EMG %>%
  group_by(Stimulus, ER_Strategy, Expectation) %>%
  summarize(
    median = median(amplitude),
    mad = mad(amplitude)
  )

kable(
  EMG_summary,
  digits = c(0, 0, 0, 5, 5),
  caption = "Descriptive statistics of EMG amplitude, separately for each condition."
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## Plot

```{r all_plot, fig.cap = "**Figure 1**. EMG amplitude in response to cue and target. **Note**: 18 values are outside the plot."}

# plot
EMG %>%
  ggplot(
    aes(
      x = ER_Strategy,
      y = amplitude,
      color = Expectation,
      fill = Expectation
    )
  ) +
  geom_pirate(
    bars = FALSE,
    cis = TRUE,
    lines = TRUE, lines_params = list(color = "black", alpha = .3),
    points = TRUE, points_params = list(color = "black", shape = 21, size = 5, alpha = .2),
    violins = TRUE, violins_params = list(size = 1),
    show.legend = TRUE
  ) +
  scale_color_viridis_d(option = "cividis") +
  scale_fill_viridis_d(option = "cividis") +
  scale_y_continuous(limits = c(-.05, .05)) +
  ggtitle("EMG") +
  facet_wrap(. ~ Stimulus) +
  theme_custom +
  theme(
    legend.position = c(.55, .8),
    axis.title.x = element_blank()
  )

```

## Analysis

Repeated measures nonparametric ANOVA (Aligned Rank Transform).

```{r all_analysis}

# repeated measures nonparametric ANOVA, Aligned Rank Transform
all_ART <- anova(
  art(amplitude ~ Stimulus * ER_Strategy * Expectation + Error(Participant),
    data = EMG
  )
) %>%
  mutate(
    eta.sq.part = `Sum Sq` / (`Sum Sq` + `Sum Sq.res`) # partial eta squared
  )

all_ART %>%
  dplyr::select(`Term`, `Df`, `Df.res`, `F` = `F value`, p.value = `Pr(>F)`, `eta.sq.part`) %>%
  kable(.,
    digits = c(0, 0, 0, 2, 3, 3),
    caption = "Repeated measures nonparametric ANOVA (Aligned Rank Transform)."
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

# Cue

Analyze EMG following cue presentation.

```{r cue_data}

# keep only EMG in response to cue
EMG_cue <- filter(EMG, Stimulus == "cue") %>%
  dplyr::select(-Stimulus)

```

## Plot

```{r cue_plot, fig.cap = "**Figure 2**. EMG amplitude in response to cue. **Note**: 9 values are outside the plot."}

# plot
EMG_cue %>%
  ggplot(
    aes(
      x = ER_Strategy,
      y = amplitude,
      color = Expectation,
      fill = Expectation
    )
  ) +
  geom_pirate(
    bars = FALSE,
    cis = TRUE,
    lines = TRUE, lines_params = list(color = "black", alpha = .3),
    points = TRUE, points_params = list(color = "black", shape = 21, size = 5, alpha = .2),
    violins = TRUE, violins_params = list(size = 1),
    show.legend = TRUE
  ) +
  scale_color_viridis_d(option = "cividis") +
  scale_fill_viridis_d(option = "cividis") +
  scale_y_continuous(limits = c(-.05, .05)) +
  ggtitle("Cue EMG") +
  theme_custom +
  theme(
    legend.position = c(.5, .15),
    axis.title.x = element_blank()
  )

```

## Analysis

Repeated measures nonparametric ANOVA (Aligned Rank Transform), post-hoc Paired Samples Wilcoxon Tests.

```{r cue_analysis}

# repeated measures nonparametric ANOVA, Aligned Rank Transform
cue_ART <- anova(
  art(amplitude ~ ER_Strategy * Expectation + Error(Participant),
    data = EMG_cue
  )
) %>%
  mutate(
    eta.sq.part = `Sum Sq` / (`Sum Sq` + `Sum Sq.res`) # partial eta squared
  )

cue_ART %>%
  dplyr::select(`Term`, `Df`, `Df.res`, `F` = `F value`, p.value = `Pr(>F)`, `eta.sq.part`) %>%
  kable(.,
    digits = c(0, 0, 0, 2, 3, 3),
    caption = "Repeated measures nonparametric ANOVA (Aligned Rank Transform)."
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

## post-hoc analyses
# Appraisal, Expected vs. Unexpected
# Paired Samples Wilcoxon Test
EMG_Appraisal_ExpectedUnexpected_Wilcoxon <-
  wilcox.test(filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Appraisal_ExpectedUnexpected_r <-
  wilcoxonPairedR(filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation %in% c("Expected", "Unexpected"))$amplitude,
    filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation %in% c("Expected", "Unexpected"))$Expectation,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Reappraisal, Expected vs. Unexpected
# Paired Samples Wilcoxon Test
EMG_Reappraisal_ExpectedUnexpected_Wilcoxon <-
  wilcox.test(filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Reappraisal_ExpectedUnexpected_r <-
  wilcoxonPairedR(filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation %in% c("Expected", "Unexpected"))$amplitude,
    filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation %in% c("Expected", "Unexpected"))$Expectation,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Expected, Appraisal vs. Reappraisal
# Paired Samples Wilcoxon Test
EMG_Expected_AppraisalReappraisal_Wilcoxon <-
  wilcox.test(filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation == "Expected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Expected_AppraisalReappraisal_r <-
  wilcoxonPairedR(filter(EMG_cue, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Expected")$amplitude,
    filter(EMG_cue, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Expected")$ER_Strategy,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Unexpected, Appraisal vs. Reappraisal
# Paired Samples Wilcoxon Test
EMG_Unexpected_AppraisalReappraisal_Wilcoxon <-
  wilcox.test(filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation == "Unexpected")$amplitude,
    filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Unexpected_AppraisalReappraisal_r <-
  wilcoxonPairedR(filter(EMG_cue, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Unexpected")$amplitude,
    filter(EMG_cue, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Unexpected")$ER_Strategy,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# summary
cue_ART_posthoc <- tibble(
  "comparison" = c(
    "Appraisal, Expected vs. Unexpected", "Reappraisal, Expected vs. Unexpected",
    "Expected, Appraisal vs. Reappraisal", "Unexpected, Appraisal vs. Reappraisal"
  ),
  "V" = c(
    EMG_Appraisal_ExpectedUnexpected_Wilcoxon$statistic, EMG_Reappraisal_ExpectedUnexpected_Wilcoxon$statistic,
    EMG_Expected_AppraisalReappraisal_Wilcoxon$statistic, EMG_Unexpected_AppraisalReappraisal_Wilcoxon$statistic
  ),
  "p" = c(
    EMG_Appraisal_ExpectedUnexpected_Wilcoxon$p.value, EMG_Reappraisal_ExpectedUnexpected_Wilcoxon$p.value,
    EMG_Expected_AppraisalReappraisal_Wilcoxon$p.value, EMG_Unexpected_AppraisalReappraisal_Wilcoxon$p.value
  ),
  "r" = c(
    EMG_Appraisal_ExpectedUnexpected_r$r, EMG_Reappraisal_ExpectedUnexpected_r$r,
    EMG_Expected_AppraisalReappraisal_r$r, EMG_Unexpected_AppraisalReappraisal_r$r
  ),
  "CI95_lower" = c(
    EMG_Appraisal_ExpectedUnexpected_r$lower.ci, EMG_Reappraisal_ExpectedUnexpected_r$lower.ci,
    EMG_Expected_AppraisalReappraisal_r$lower.ci, EMG_Unexpected_AppraisalReappraisal_r$lower.ci
  ),
  "CI95_upper" = c(
    EMG_Appraisal_ExpectedUnexpected_r$upper.ci, EMG_Reappraisal_ExpectedUnexpected_r$upper.ci,
    EMG_Expected_AppraisalReappraisal_r$upper.ci, EMG_Unexpected_AppraisalReappraisal_r$upper.ci
  )
) %>%
  mutate("p" = p.adjust(p, method = "holm")) # Bonferroni-Holm correction on post-hoc tests

kable(cue_ART_posthoc,
  digits = c(0, 0, 3, 2, 2, 2),
  caption = "Post-hoc Repeated Samples Wilcoxon tests (Bonferroni-Holm p-value correction), bootstrapped effect size."
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

# Target

Analyze EMG following target presentation.

```{r target_data}

# keep only EMG in response to target
EMG_target <- filter(EMG, Stimulus == "target") %>%
  dplyr::select(-Stimulus)

```

## Plot

```{r target_plot, fig.cap = "**Figure 3.** EMG amplitude in response to target. **Note**: 9 values are outside the plot."}

# plot
EMG_target %>%
  ggplot(
    aes(
      x = ER_Strategy,
      y = amplitude,
      color = Expectation,
      fill = Expectation
    )
  ) +
  geom_pirate(
    bars = FALSE,
    cis = TRUE,
    lines = TRUE, lines_params = list(color = "black", alpha = .3),
    points = TRUE, points_params = list(color = "black", shape = 21, size = 5, alpha = .2),
    violins = TRUE, violins_params = list(size = 1),
    show.legend = TRUE
  ) +
  scale_color_viridis_d(option = "cividis") +
  scale_fill_viridis_d(option = "cividis") +
  scale_y_continuous(limits = c(-.05, .05)) +
  ggtitle("Target EMG") +
  theme_custom +
  theme(
    legend.position = c(.5, .15),
    axis.title.x = element_blank()
  )

```

## Analysis

Repeated measures nonparametric ANOVA (Aligned Rank Transform), post-hoc Paired Samples Wilcoxon Tests.

```{r target_analysis}

# repeated measures nonparametric ANOVA, Aligned Rank Transform
target_ART <- anova(
  art(amplitude ~ ER_Strategy * Expectation + Error(Participant),
    data = EMG_target
  )
) %>%
  mutate(
    eta.sq.part = `Sum Sq` / (`Sum Sq` + `Sum Sq.res`) # partial eta squared
  )

target_ART %>%
  dplyr::select(`Term`, `Df`, `Df.res`, `F` = `F value`, p.value = `Pr(>F)`, `eta.sq.part`) %>%
  kable(.,
    digits = c(0, 0, 0, 2, 3, 3),
    caption = "Repeated measures nonparametric ANOVA (Aligned Rank Transform)."
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

## post-hoc analyses
# Appraisal, Expected vs. Unexpected
# Paired Samples Wilcoxon Test
EMG_Appraisal_ExpectedUnexpected_Wilcoxon <-
  wilcox.test(filter(EMG_target, ER_Strategy == "Appraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_target, ER_Strategy == "Appraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Appraisal_ExpectedUnexpected_r <-
  wilcoxonPairedR(filter(EMG_target, ER_Strategy == "Appraisal" & Expectation %in% c("Expected", "Unexpected"))$amplitude,
    filter(EMG_target, ER_Strategy == "Appraisal" & Expectation %in% c("Expected", "Unexpected"))$Expectation,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Reappraisal, Expected vs. Unexpected
# Paired Samples Wilcoxon Test
EMG_Reappraisal_ExpectedUnexpected_Wilcoxon <-
  wilcox.test(filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Reappraisal_ExpectedUnexpected_r <-
  wilcoxonPairedR(filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation %in% c("Expected", "Unexpected"))$amplitude,
    filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation %in% c("Expected", "Unexpected"))$Expectation,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Expected, Appraisal vs. Reappraisal
# Paired Samples Wilcoxon Test
EMG_Expected_AppraisalReappraisal_Wilcoxon <-
  wilcox.test(filter(EMG_target, ER_Strategy == "Appraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation == "Expected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Expected_AppraisalReappraisal_r <-
  wilcoxonPairedR(filter(EMG_target, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Expected")$amplitude,
    filter(EMG_target, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Expected")$ER_Strategy,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Unexpected, Appraisal vs. Reappraisal
# Paired Samples Wilcoxon Test
EMG_Unexpected_AppraisalReappraisal_Wilcoxon <-
  wilcox.test(filter(EMG_target, ER_Strategy == "Appraisal" & Expectation == "Unexpected")$amplitude,
    filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Unexpected_AppraisalReappraisal_r <-
  wilcoxonPairedR(filter(EMG_target, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Unexpected")$amplitude,
    filter(EMG_target, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Unexpected")$ER_Strategy,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# summary
target_ART_posthoc <- tibble(
  "comparison" = c(
    "Appraisal, Expected vs. Unexpected", "Reappraisal, Expected vs. Unexpected",
    "Expected, Appraisal vs. Reappraisal", "Unexpected, Appraisal vs. Reappraisal"
  ),
  "V" = c(
    EMG_Appraisal_ExpectedUnexpected_Wilcoxon$statistic, EMG_Reappraisal_ExpectedUnexpected_Wilcoxon$statistic,
    EMG_Expected_AppraisalReappraisal_Wilcoxon$statistic, EMG_Unexpected_AppraisalReappraisal_Wilcoxon$statistic
  ),
  "p" = c(
    EMG_Appraisal_ExpectedUnexpected_Wilcoxon$p.value, EMG_Reappraisal_ExpectedUnexpected_Wilcoxon$p.value,
    EMG_Expected_AppraisalReappraisal_Wilcoxon$p.value, EMG_Unexpected_AppraisalReappraisal_Wilcoxon$p.value
  ),
  "r" = c(
    EMG_Appraisal_ExpectedUnexpected_r$r, EMG_Reappraisal_ExpectedUnexpected_r$r,
    EMG_Expected_AppraisalReappraisal_r$r, EMG_Unexpected_AppraisalReappraisal_r$r
  ),
  "CI95_lower" = c(
    EMG_Appraisal_ExpectedUnexpected_r$lower.ci, EMG_Reappraisal_ExpectedUnexpected_r$lower.ci,
    EMG_Expected_AppraisalReappraisal_r$lower.ci, EMG_Unexpected_AppraisalReappraisal_r$lower.ci
  ),
  "CI95_upper" = c(
    EMG_Appraisal_ExpectedUnexpected_r$upper.ci, EMG_Reappraisal_ExpectedUnexpected_r$upper.ci,
    EMG_Expected_AppraisalReappraisal_r$upper.ci, EMG_Unexpected_AppraisalReappraisal_r$upper.ci
  )
) %>%
  mutate("p" = p.adjust(p, method = "holm")) # Bonferroni-Holm correction on post-hoc tests

kable(target_ART_posthoc,
  digits = c(0, 0, 3, 2, 2, 2),
  caption = "Post-hoc Repeated Samples Wilcoxon tests (Bonferroni-Holm p-value correction), bootstrapped effect size."
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

# Exploratory Analysis

## Block order

Participants with even numbers started with reappraisal blocks, participants with uneven numbers started with appraisal block. Does block order influence EMG amplitude?

```{r block_data}

EMG <- EMG %>%
  mutate(
    start_block = as_factor(
      ifelse(
        as.numeric(Participant) %% 2 == 0,
        "start_Reappraisal",
        "start_Appraisal"
      )
    )
  )

```

```{r block_plots, fig.cap = "**Figure 4**. EMG amplitude in response to cue and target, separately for participants who started with Appraisal vs. Reappraisal blocks. **Note**: 9 values are outside the plot."}

# plot start appraisal
plot_start_appraisal <-
  EMG %>%
  filter(start_block == "start_Appraisal") %>%
  ggplot(
    aes(
      x = ER_Strategy,
      y = amplitude,
      color = Expectation,
      fill = Expectation
    )
  ) +
  geom_pirate(
    bars = FALSE,
    cis = TRUE,
    lines = TRUE, lines_params = list(color = "black", alpha = .3),
    points = TRUE, points_params = list(color = "black", shape = 21, size = 5, alpha = .2),
    violins = TRUE, violins_params = list(size = 1),
    show.legend = TRUE
  ) +
  scale_color_viridis_d(option = "cividis") +
  scale_fill_viridis_d(option = "cividis") +
  scale_y_continuous(limits = c(-.05, .05)) +
  ggtitle("appraisal first") +
  facet_wrap(. ~ Stimulus) +
  theme_custom +
  theme(
    legend.position = c(.55, .8),
    axis.title.x = element_blank()
  )

# plot start reappraisal
plot_start_reappraisal <-
  EMG %>%
  filter(start_block == "start_Reappraisal") %>%
  ggplot(
    aes(
      x = ER_Strategy,
      y = amplitude,
      color = Expectation,
      fill = Expectation
    )
  ) +
  geom_pirate(
    bars = FALSE,
    cis = TRUE,
    lines = TRUE, lines_params = list(color = "black", alpha = .3),
    points = TRUE, points_params = list(color = "black", shape = 21, size = 5, alpha = .2),
    violins = TRUE, violins_params = list(size = 1),
    show.legend = TRUE
  ) +
  scale_color_viridis_d(option = "cividis") +
  scale_fill_viridis_d(option = "cividis") +
  scale_y_continuous(limits = c(-.05, .05)) +
  ggtitle("reappraisal first") +
  facet_wrap(. ~ Stimulus) +
  theme_custom +
  theme(
    legend.position = c(.45, .2),
    axis.title.x = element_blank()
  )

# arrange plots in two columns
plot_block_data <- plot_grid(
  plot_start_appraisal + theme(legend.position = "none"),
  plot_start_reappraisal + theme(legend.position = c(.5, .9)),
  align = "vh",
  hjust = -1,
  nrow = 2
)

plot_block_data

```

The package `ARTool` does not handle split-plot designs, therefore we will run non-parametric permutation tests via the function `ezPerm` from the package `ez`.

```{r block_analysis, eval = FALSE}

block_ez <-
  ezPerm(
    data = EMG,
    dv = amplitude,
    wid = Participant,
    within = c(Stimulus, ER_Strategy, Expectation),
    between = start_block,
    perms = 1000
  )

saveRDS(block_ez, file = paste0(getwd(), "/block_ez.rds"), compress = "gzip")

```

```{r block_analysis_results}

block_ez <- readRDS(paste0(getwd(), "/block_ez.rds"))

block_ez %>%
  kable(.,
    digits = c(0, 3, 0),
    caption = "Split-plot ANOVA, control for block order (Appraisal first or Reappraisal first)."
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

Results are **not** influenced by block order.

##  Correlation

Reviewer's comment:

>It seems as though the primary difference during the target phase is for unexpected negative feedback, and it is reduced for reappraised vs. appraised trials. Does something going on during anticipation predict this effect? For example, does the difference in reappraise vs. appraise during anticipation of unexpected feedback (or something else during anticipation) correlate/predict this effect?

```{r spearman}

EMG_Unexpected_corr <- EMG %>%
  filter(
    Expectation == "Unexpected"
  ) %>%
  dplyr::select(Participant, Stimulus, ER_Strategy, amplitude) %>%
  pivot_wider(
    id_cols = Participant,
    names_from = c(Stimulus, ER_Strategy),
    values_from = amplitude
  ) %>%
  mutate(
    cue_ReappraisalVsAppraisal = cue_Reappraisal - cue_Appraisal,
    target_ReappraisalVsAppraisal = target_Reappraisal - target_Appraisal
  ) %>%
  dplyr::select(Participant, cue_ReappraisalVsAppraisal, target_ReappraisalVsAppraisal)

# Spearman's correlation
corr <- cor.test(
  x = pull(EMG_Unexpected_corr, "cue_ReappraisalVsAppraisal"),
  y = pull(EMG_Unexpected_corr, "target_ReappraisalVsAppraisal"),
  method = "spearman",
  alternative = "two.sided",
  exact = TRUE,
  continuity = TRUE
)

tibble(
  "rho" = corr$statistic,
  "p-value" = corr$p.value
) %>%
  kable(.,
    digits = c(0, 3),
    caption = ""
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

The difference between Appraisal and Reappraisal during anticipation does **not** statistically correlate with the same difference during the target phase.

***
***

# Session Info

```{r}

sessionInfo()

```
