---
title: "Analysis EMG (no outliers)"
author: '[Antonio Schettino](https://osf.io/zbv65/ "Antonio Schettino")'
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    highlight: tango
    theme: united
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

# ### install packages
# install.packages(c(
#   "knitr", "kableExtra",
#   "tidyverse",
#   "rcompanion", "ARTool", "ez"
#   "viridis", "cowplot"
# ),
# dependencies = TRUE
# )
# devtools::install_github("mikabr/ggpirate", dependencies = TRUE)

### load packages
library(knitr)
library(kableExtra)
library(tidyverse)
library(rcompanion)
library(ARTool)
library(ez)
library(viridis)
library(ggpirate)
library(cowplot)

# for RMarkdown
options(digits = 2) # number of decimal digits
opts_chunk$set(
  fig.dim = c(10, 10),
  warning = FALSE, # no package warnings
  message = FALSE # no package messages
)

# ggplot custom theme
theme_custom <- theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(
      hjust = .5,
      size = 20
    ),
    legend.title = element_blank(),
    plot.title = element_text(size = 26, hjust = .5),
    legend.box.background = element_rect(color = "transparent") # transparent outer line
  )

seed_smorfia <- 1 # Italia!
set.seed(seed_smorfia) # seed for random number generation

```

***
***

# Participants who did not believe the cover story

Eleven participants did not believe the experimental manipulation, and the responses of 3 participants are missing. Here we confirm that the results are qualitatively similar with and without these 14 outliers.

```{r all_data}

EMG <- read_csv("EMG.csv") %>%
  pivot_longer(
    Reappraisal_Unexpected:Appraisal_Expected,
    names_to = "condition",
    values_to = "amplitude"
  ) %>%
  separate(
    condition,
    c("ER_Strategy", "Expectation"),
    sep = "_",
    remove = TRUE
  ) %>%
  mutate(
    Participant = as_factor(Participant),
    Believed_Story = as_factor(Believed_Story),
    Stimulus = as_factor(Stimulus),
    ER_Strategy = as_factor(ER_Strategy),
    Expectation = as_factor(Expectation)
  ) %>%
  filter(Believed_Story == "yes")

# summary
EMG_summary <-
  EMG %>%
  group_by(Stimulus, ER_Strategy, Expectation) %>%
  summarize(
    median = median(amplitude),
    mad = mad(amplitude)
  )

kable(
  EMG_summary,
  digits = c(0, 0, 0, 5, 5),
  caption = "Descriptive statistics of EMG amplitude, separately for each condition."
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## Plot

```{r all_plot, fig.cap = "**Figure 1**. EMG amplitude in response to cue and target. **Note**: 18 values are outside the plot."}

# plot
EMG %>%
  ggplot(
    aes(
      x = ER_Strategy,
      y = amplitude,
      color = Expectation,
      fill = Expectation
    )
  ) +
  geom_pirate(
    bars = FALSE,
    cis = TRUE,
    lines = TRUE, lines_params = list(color = "black", alpha = .3),
    points = TRUE, points_params = list(color = "black", shape = 21, size = 5, alpha = .2),
    violins = TRUE, violins_params = list(size = 1),
    show.legend = TRUE
  ) +
  scale_color_viridis_d(option = "cividis") +
  scale_fill_viridis_d(option = "cividis") +
  scale_y_continuous(limits = c(-.05, .05)) +
  ggtitle("EMG") +
  facet_wrap(. ~ Stimulus) +
  theme_custom +
  theme(
    legend.position = c(.55, .8),
    axis.title.x = element_blank()
  )

```

## Analysis

```{r all_analysis}

# repeated measures nonparametric ANOVA, Aligned Rank Transform
all_ART <- anova(
  art(amplitude ~ Stimulus * ER_Strategy * Expectation + Error(Participant),
    data = EMG
  )
) %>%
  mutate(
    eta.sq.part = `Sum Sq` / (`Sum Sq` + `Sum Sq.res`) # partial eta squared
  )

all_ART %>%
  dplyr::select(`Term`, `Df`, `Df.res`, `F` = `F value`, p.value = `Pr(>F)`, `eta.sq.part`) %>%
  kable(.,
    digits = c(0, 0, 0, 2, 3, 3),
    caption = "Repeated measures nonparametric ANOVA (Aligned Rank Transform)."
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

# Cue

Analyze EMG following cue presentation.

```{r cue_data}

# keep only EMG in response to cue
EMG_cue <- filter(EMG, Stimulus == "cue") %>%
  dplyr::select(-Stimulus)

```

## Plot

```{r cue_plot, fig.cap = "**Figure 2**. EMG amplitude in response to cue. **Note**: 9 values are outside the plot."}

# plot
EMG_cue %>%
  ggplot(
    aes(
      x = ER_Strategy,
      y = amplitude,
      color = Expectation,
      fill = Expectation
    )
  ) +
  geom_pirate(
    bars = FALSE,
    cis = TRUE,
    lines = TRUE, lines_params = list(color = "black", alpha = .3),
    points = TRUE, points_params = list(color = "black", shape = 21, size = 5, alpha = .2),
    violins = TRUE, violins_params = list(size = 1),
    show.legend = TRUE
  ) +
  scale_color_viridis_d(option = "cividis") +
  scale_fill_viridis_d(option = "cividis") +
  scale_y_continuous(limits = c(-.05, .05)) +
  ggtitle("Cue EMG") +
  theme_custom +
  theme(
    legend.position = c(.5, .15),
    axis.title.x = element_blank()
  )

```

## Analysis

Repeated measures nonparametric ANOVA (Aligned Rank Transform), post-hoc Paired Samples Wilcoxon Tests.

```{r cue_analysis}

# repeated measures nonparametric ANOVA, Aligned Rank Transform
cue_ART <- anova(
  art(amplitude ~ ER_Strategy * Expectation + Error(Participant),
    data = EMG_cue
  )
) %>%
  mutate(
    eta.sq.part = `Sum Sq` / (`Sum Sq` + `Sum Sq.res`) # partial eta squared
  )

cue_ART %>%
  dplyr::select(`Term`, `Df`, `Df.res`, `F` = `F value`, p.value = `Pr(>F)`, `eta.sq.part`) %>%
  kable(.,
    digits = c(0, 0, 0, 2, 3, 3),
    caption = "Repeated measures nonparametric ANOVA (Aligned Rank Transform)."
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

## post-hoc analyses
# Appraisal, Expected vs. Unexpected
# Paired Samples Wilcoxon Test
EMG_Appraisal_ExpectedUnexpected_Wilcoxon <-
  wilcox.test(filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Appraisal_ExpectedUnexpected_r <-
  wilcoxonPairedR(filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation %in% c("Expected", "Unexpected"))$amplitude,
    filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation %in% c("Expected", "Unexpected"))$Expectation,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Reappraisal, Expected vs. Unexpected
# Paired Samples Wilcoxon Test
EMG_Reappraisal_ExpectedUnexpected_Wilcoxon <-
  wilcox.test(filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Reappraisal_ExpectedUnexpected_r <-
  wilcoxonPairedR(filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation %in% c("Expected", "Unexpected"))$amplitude,
    filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation %in% c("Expected", "Unexpected"))$Expectation,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Expected, Appraisal vs. Reappraisal
# Paired Samples Wilcoxon Test
EMG_Expected_AppraisalReappraisal_Wilcoxon <-
  wilcox.test(filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation == "Expected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Expected_AppraisalReappraisal_r <-
  wilcoxonPairedR(filter(EMG_cue, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Expected")$amplitude,
    filter(EMG_cue, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Expected")$ER_Strategy,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Unexpected, Appraisal vs. Reappraisal
# Paired Samples Wilcoxon Test
EMG_Unexpected_AppraisalReappraisal_Wilcoxon <-
  wilcox.test(filter(EMG_cue, ER_Strategy == "Appraisal" & Expectation == "Unexpected")$amplitude,
    filter(EMG_cue, ER_Strategy == "Reappraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Unexpected_AppraisalReappraisal_r <-
  wilcoxonPairedR(filter(EMG_cue, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Unexpected")$amplitude,
    filter(EMG_cue, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Unexpected")$ER_Strategy,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# summary
cue_ART_posthoc <- tibble(
  "comparison" = c(
    "Appraisal, Expected vs. Unexpected", "Reappraisal, Expected vs. Unexpected",
    "Expected, Appraisal vs. Reappraisal", "Unexpected, Appraisal vs. Reappraisal"
  ),
  "V" = c(
    EMG_Appraisal_ExpectedUnexpected_Wilcoxon$statistic, EMG_Reappraisal_ExpectedUnexpected_Wilcoxon$statistic,
    EMG_Expected_AppraisalReappraisal_Wilcoxon$statistic, EMG_Unexpected_AppraisalReappraisal_Wilcoxon$statistic
  ),
  "p" = c(
    EMG_Appraisal_ExpectedUnexpected_Wilcoxon$p.value, EMG_Reappraisal_ExpectedUnexpected_Wilcoxon$p.value,
    EMG_Expected_AppraisalReappraisal_Wilcoxon$p.value, EMG_Unexpected_AppraisalReappraisal_Wilcoxon$p.value
  ),
  "r" = c(
    EMG_Appraisal_ExpectedUnexpected_r$r, EMG_Reappraisal_ExpectedUnexpected_r$r,
    EMG_Expected_AppraisalReappraisal_r$r, EMG_Unexpected_AppraisalReappraisal_r$r
  ),
  "CI95_lower" = c(
    EMG_Appraisal_ExpectedUnexpected_r$lower.ci, EMG_Reappraisal_ExpectedUnexpected_r$lower.ci,
    EMG_Expected_AppraisalReappraisal_r$lower.ci, EMG_Unexpected_AppraisalReappraisal_r$lower.ci
  ),
  "CI95_upper" = c(
    EMG_Appraisal_ExpectedUnexpected_r$upper.ci, EMG_Reappraisal_ExpectedUnexpected_r$upper.ci,
    EMG_Expected_AppraisalReappraisal_r$upper.ci, EMG_Unexpected_AppraisalReappraisal_r$upper.ci
  )
) %>%
  mutate("p" = p.adjust(p, method = "holm")) # Bonferroni-Holm correction on post-hoc tests

kable(cue_ART_posthoc,
  digits = c(0, 0, 3, 2, 2, 2),
  caption = "Post-hoc Repeated Samples Wilcoxon tests (Bonferroni-Holm p-value correction), bootstrapped effect size."
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

# Target

Analyze EMG following target presentation.

```{r target_data}

# keep only EMG in response to target
EMG_target <- filter(EMG, Stimulus == "target") %>%
  dplyr::select(-Stimulus)

```

## Plot

```{r target_plot, fig.cap = "**Figure 3**. EMG amplitude in response to target. **Note**: 9 values are outside the plot."}

# plot
EMG_target %>%
  ggplot(
    aes(
      x = ER_Strategy,
      y = amplitude,
      color = Expectation,
      fill = Expectation
    )
  ) +
  geom_pirate(
    bars = FALSE,
    cis = TRUE,
    lines = TRUE, lines_params = list(color = "black", alpha = .3),
    points = TRUE, points_params = list(color = "black", shape = 21, size = 5, alpha = .2),
    violins = TRUE, violins_params = list(size = 1),
    show.legend = TRUE
  ) +
  scale_color_viridis_d(option = "cividis") +
  scale_fill_viridis_d(option = "cividis") +
  scale_y_continuous(limits = c(-.05, .05)) +
  ggtitle("Target EMG") +
  theme_custom +
  theme(
    legend.position = c(.5, .15),
    axis.title.x = element_blank()
  )

```

## Analysis

Repeated measures nonparametric ANOVA (Aligned Rank Transform), post-hoc Paired Samples Wilcoxon Tests.

```{r target_analysis}

# repeated measures nonparametric ANOVA, Aligned Rank Transform
target_ART <- anova(
  art(amplitude ~ ER_Strategy * Expectation + Error(Participant),
    data = EMG_target
  )
) %>%
  mutate(
    eta.sq.part = `Sum Sq` / (`Sum Sq` + `Sum Sq.res`) # partial eta squared
  )

target_ART %>%
  dplyr::select(`Term`, `Df`, `Df.res`, `F` = `F value`, p.value = `Pr(>F)`, `eta.sq.part`) %>%
  kable(.,
    digits = c(0, 0, 0, 2, 3, 3),
    caption = "Repeated measures nonparametric ANOVA (Aligned Rank Transform)."
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

## post-hoc analyses
# Appraisal, Expected vs. Unexpected
# Paired Samples Wilcoxon Test
EMG_Appraisal_ExpectedUnexpected_Wilcoxon <-
  wilcox.test(filter(EMG_target, ER_Strategy == "Appraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_target, ER_Strategy == "Appraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Appraisal_ExpectedUnexpected_r <-
  wilcoxonPairedR(filter(EMG_target, ER_Strategy == "Appraisal" & Expectation %in% c("Expected", "Unexpected"))$amplitude,
    filter(EMG_target, ER_Strategy == "Appraisal" & Expectation %in% c("Expected", "Unexpected"))$Expectation,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Reappraisal, Expected vs. Unexpected
# Paired Samples Wilcoxon Test
EMG_Reappraisal_ExpectedUnexpected_Wilcoxon <-
  wilcox.test(filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Reappraisal_ExpectedUnexpected_r <-
  wilcoxonPairedR(filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation %in% c("Expected", "Unexpected"))$amplitude,
    filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation %in% c("Expected", "Unexpected"))$Expectation,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Expected, Appraisal vs. Reappraisal
# Paired Samples Wilcoxon Test
EMG_Expected_AppraisalReappraisal_Wilcoxon <-
  wilcox.test(filter(EMG_target, ER_Strategy == "Appraisal" & Expectation == "Expected")$amplitude,
    filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation == "Expected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Expected_AppraisalReappraisal_r <-
  wilcoxonPairedR(filter(EMG_target, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Expected")$amplitude,
    filter(EMG_target, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Expected")$ER_Strategy,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# Unexpected, Appraisal vs. Reappraisal
# Paired Samples Wilcoxon Test
EMG_Unexpected_AppraisalReappraisal_Wilcoxon <-
  wilcox.test(filter(EMG_target, ER_Strategy == "Appraisal" & Expectation == "Unexpected")$amplitude,
    filter(EMG_target, ER_Strategy == "Reappraisal" & Expectation == "Unexpected")$amplitude,
    alternative = "two.sided",
    mu = 0,
    paired = TRUE,
    exact = TRUE,
    correct = TRUE,
    conf.int = TRUE, conf.level = 0.95
  )

# effect size r
EMG_Unexpected_AppraisalReappraisal_r <-
  wilcoxonPairedR(filter(EMG_target, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Unexpected")$amplitude,
    filter(EMG_target, ER_Strategy %in% c("Appraisal", "Reappraisal") & Expectation == "Unexpected")$ER_Strategy,
    ci = TRUE, conf = 0.95,
    type = "bca", R = 5000,
    cases = TRUE
  )

# summary
target_ART_posthoc <- tibble(
  "comparison" = c(
    "Appraisal, Expected vs. Unexpected", "Reappraisal, Expected vs. Unexpected",
    "Expected, Appraisal vs. Reappraisal", "Unexpected, Appraisal vs. Reappraisal"
  ),
  "V" = c(
    EMG_Appraisal_ExpectedUnexpected_Wilcoxon$statistic, EMG_Reappraisal_ExpectedUnexpected_Wilcoxon$statistic,
    EMG_Expected_AppraisalReappraisal_Wilcoxon$statistic, EMG_Unexpected_AppraisalReappraisal_Wilcoxon$statistic
  ),
  "p" = c(
    EMG_Appraisal_ExpectedUnexpected_Wilcoxon$p.value, EMG_Reappraisal_ExpectedUnexpected_Wilcoxon$p.value,
    EMG_Expected_AppraisalReappraisal_Wilcoxon$p.value, EMG_Unexpected_AppraisalReappraisal_Wilcoxon$p.value
  ),
  "r" = c(
    EMG_Appraisal_ExpectedUnexpected_r$r, EMG_Reappraisal_ExpectedUnexpected_r$r,
    EMG_Expected_AppraisalReappraisal_r$r, EMG_Unexpected_AppraisalReappraisal_r$r
  ),
  "CI95_lower" = c(
    EMG_Appraisal_ExpectedUnexpected_r$lower.ci, EMG_Reappraisal_ExpectedUnexpected_r$lower.ci,
    EMG_Expected_AppraisalReappraisal_r$lower.ci, EMG_Unexpected_AppraisalReappraisal_r$lower.ci
  ),
  "CI95_upper" = c(
    EMG_Appraisal_ExpectedUnexpected_r$upper.ci, EMG_Reappraisal_ExpectedUnexpected_r$upper.ci,
    EMG_Expected_AppraisalReappraisal_r$upper.ci, EMG_Unexpected_AppraisalReappraisal_r$upper.ci
  )
) %>%
  mutate("p" = p.adjust(p, method = "holm")) # Bonferroni-Holm correction on post-hoc tests

kable(target_ART_posthoc,
  digits = c(0, 0, 3, 2, 2, 2),
  caption = "Post-hoc Repeated Samples Wilcoxon tests (Bonferroni-Holm p-value correction), bootstrapped effect size."
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

***
***

# Session Info

```{r}

sessionInfo()

```
